Childhood exposure to non-persistent endocrine disruptors, glucocorticosteroids, and attentional function: A cross-sectional study based on the parametric g-formula

### 2024-07-19 output/paper/_figures.qmd
Comments AJE.
Working on comments Reviewer 1 points 9 to 11.

<!-- title: "Figures" -->

<!-- format: -->
<!--   docx: -->
<!--     reference-doc: reference_document.docx -->
<!--   pdf: -->
<!--     pdf-engine: lualatex -->
<!--     number-sections: true -->
<!--     toc: false -->
<!--     include-in-header: -->
<!--       text: | -->
<!--         \usepackage{lineno} -->
<!--         \usepackage{rotating} -->
<!--         \usepackage{float} -->
<!--         \usepackage{typearea} -->
<!--     include-before-body: -->
<!--       text: | -->
<!--         \linenumbers -->

<!-- link-citations: true -->
<!-- linkcolor: blue -->

<!-- editor_options: -->
<!--   chunk_output_type: console -->

<!-- execute: -->
<!--   echo: false -->
<!--   warning: false -->
<!--   error: false -->
<!--   message: false -->

@@

>>>>

---
editor_options:
  chunk_output_type: console
---

@@

# DAG

::: {#fig-dag}
```{dot}
//| fig-width: 9
//| fig-height: 10

digraph simple {
    graph [rankdir=TB]

    node [shape=ellipse]
    U [label='Confounders']

    node [shape=box, height=0.3, width=0.3]
    E [label='EDCs']
    M [label='Glucocorticosteroids']
    Y [label='HRT-SE']
    { rank = same; E M}

    E -> M [minlen=3]
    E -> Y
    M -> Y
    U -> E
    U -> M
    U -> Y
  }
```
**Simplified {{< acr dag first_use=true >}} showing {{< acr edcs first_use=true >}}, {{< acr hitrtse first_use=true >}}, and the glucocorticosteroids.**
:::

### 2024-07-19 output/paper/_si.qmd
Comments AJE.
Working on comments Reviewer 1 points 9 to 11.

```{r}
targets::tar_load(
  dplyr::contains("codebook_"),
  store = store
)
```

::: {#supptbl-codebook-1}
```{r codebook-1}
codebook_1
```
**Codebook for the covariates used in the estimation of the marginal comparisons of {{< acr edcs first_use=true >}} on {{< acr hitrtse first_use=true >}}.**
:::

::: {#supptbl-codebook-2}
```{r codebook-2}
codebook_2
```
**Codebook for the covariates used in the estimation of the marginal comparisons of {{< acr edcs first_use=true >}} on the glucocorticosteroids.**
:::

::: {#supptbl-codebook-3}
```{r codebook-3}
codebook_3
```
**Codebook for the covariates used in the estimation of the marginal comparisons of the glucocorticosteroids on {{< acr hitrtse first_use=true >}}.**
:::

@@

@@

@@

tidy_res_meffects(
  df = marginal_comparisons_1SA,
  sa_var = "e3_sex",
  outcome = NULL,
  which_res = "comparisons",
  num_digits_est = num_digits_est,
  num_digits_sig = num_digits_sig
)$plot
```
**Marginal contrasts for effect modification by sex of a decrease from the 90th to the 10th percentile of the {{< acr edcs first_use=true >}} on {{< acr hitrtse first_use=true >}} expressed in {{< acr mss >}} (HELIX subcohort; 2013-2016).** 
Circles and triangles indicate effect estimates. Solid lines indicate the $95\%$ {{< acr ci >}}. The size of the circles represents the *S* value of the effect estimate [@RafiGreenland:2020].
:::

@@

tidy_res_meffects(
  df = marginal_comparisons_2SA |>
    tidylog::filter(
      !grepl("metabolism", outcome)
    ),
  sa_var = "e3_sex",
  outcome = NULL,
  which_res = "comparisons",
  num_digits_est = num_digits_est,
  num_digits_sig = num_digits_sig
)$plot
```
**Marginal contrasts for effect modification by sex of a decrease from the 90th to the 10th percentile of the {{< acr edcs first_use=true >}} on the glucocorticosteroids expressed in ng/ml (HELIX subcohort; 2013-2016).** 
Circles and triangles indicate effect estimates. Solid lines indicate the $95\%$ {{< acr ci >}}. The size of the circles represents the *S* value of the effect estimate [@RafiGreenland:2020].
:::

@@

tidy_res_meffects(
  df = marginal_comparisons_3SA,
  sa_var = "e3_sex",
  outcome = NULL,
  which_res = "comparisons",
  num_digits_est = num_digits_est,
  num_digits_sig = num_digits_sig
)$plot
```
**Marginal contrasts for effect modification by sex of a decrease from the 90th to the 10th percentile of the glucocorticosteroids on {{< acr hitrtse first_use=true >}} expressed in {{< acr mss >}} (HELIX subcohort; 2013-2016).** 
Circles and triangles indicate effect estimates. Solid lines indicate the $95\%$ {{< acr ci >}}. The size of the circles represents the *S* value of the effect estimate [@RafiGreenland:2020]. Abbreviations: cortisone production (cortisone prod.); cortisol production (cortisol prod.); corticost. prod. (corticosterone production).
:::

>>>>

```{r}
targets::tar_load(
  "codebooks",
  store = store
)
```

::: {#supptbl-codebooks}
```{r codebooks}
codebooks
```
**Codebook for the covariates used in the estimation of the marginal comparisons of {{< acr edcs first_use=true >}} on {{< acr hitrtse first_use=true >}} (`q1`), {{< acr edcs first_use=true >}} on the glucocorticosteroids (`q2`), and glucocorticosteroids on {{< acr hitrtse first_use=true >}} (`q3`).**
:::

@@

### Concentrations of the endocrine disruptors

::: {#supptbl-chems-desc}
```{r chems-desc}
targets::tar_load("desc_chems_mets", store = store)
desc_chems_mets$desc_chems_full
```
**Participants non-persistent {{< acr edcs first_use=true >}} concentrations, by cohort and overall (HELIX subcohort; 2013-2016).**
:::

@@

::: {#supptbl-marginal-1sa}
```{r marginal-1sa-tbl}
m1sa <- tidy_res_meffects(
  df = marginal_comparisons_1SA,
  sa_var = "e3_sex",
  outcome = NULL,
  which_res = "comparisons",
  num_digits_est = num_digits_est,
  num_digits_sig = num_digits_sig
)

m1sa$df_gt
```
**Marginal contrasts for effect modification by sex of a increase from the 10th to the 90th percentile of the {{< acr edcs first_use=true >}} on {{< acr hitrtse first_use=true >}} expressed in {{< acr mss >}} (HELIX subcohort; 2013-2016).**
:::

::: {#supptbl-marginal-2sa}
```{r marginal-2sa-tbl}
m2sa <- tidy_res_meffects(
  df = marginal_comparisons_2SA |>
    tidylog::filter(
      !grepl("metabolism", outcome)
    ),
  sa_var = "e3_sex",
  outcome = NULL,
  which_res = "comparisons",
  num_digits_est = num_digits_est,
  num_digits_sig = num_digits_sig
)

m2sa$df_gt
```
**Marginal contrasts for effect modification by sex of a increase from the 10th to the 90th percentile of the {{< acr edcs first_use=true >}} on the glucocorticosteroids expressed in ng/ml (HELIX subcohort; 2013-2016).**
:::

::: {#supptbl-marginal-3sa}
```{r marginal-3sa-tbl}
m3sa <- tidy_res_meffects(
  df = marginal_comparisons_3SA,
  sa_var = "e3_sex",
  outcome = NULL,
  which_res = "comparisons",
  num_digits_est = num_digits_est,
  num_digits_sig = num_digits_sig
)

m3sa$df_gt
```
**Marginal contrasts for effect modification by sex of a increase from the 10th to the 90th percentile of the glucocorticosteroids on {{< acr hitrtse first_use=true >}} expressed in {{< acr mss >}} (HELIX subcohort; 2013-2016).**
:::

@@

m1sa$plot
```
**Marginal contrasts on the logarithmic scale for effect modification by sex of a increase from the 10th to the 90th percentile of the {{< acr edcs first_use=true >}} on {{< acr hitrtse first_use=true >}} expressed in {{< acr mss >}} (HELIX subcohort; 2013-2016).** 
Circles and triangles indicate effect estimates. Solid lines indicate the $95\%$ {{< acr ci >}}. The size of the circles represents the *S* value of the effect estimate [@RafiGreenland:2020].
:::

@@

m2sa$plot
```
**Marginal contrasts on the logarithmic scale for effect modification by sex of a increase from the 10th to the 90th percentile of the {{< acr edcs first_use=true >}} on the glucocorticosteroids expressed in ng/ml (HELIX subcohort; 2013-2016).** 
Circles and triangles indicate effect estimates. Solid lines indicate the $95\%$ {{< acr ci >}}. The size of the circles represents the *S* value of the effect estimate [@RafiGreenland:2020].
:::

@@

m3sa$plot
```
**Marginal contrasts on the logarithmic scale for effect modification by sex of a increase from the 10th to the 90th percentile of the glucocorticosteroids on {{< acr hitrtse first_use=true >}} expressed in {{< acr mss >}} (HELIX subcohort; 2013-2016).** 
Circles and triangles indicate effect estimates. Solid lines indicate the $95\%$ {{< acr ci >}}. The size of the circles represents the *S* value of the effect estimate [@RafiGreenland:2020]. Abbreviations: cortisone production (cortisone prod.); cortisol production (cortisol prod.); corticost. prod. (corticosterone production).
:::

### 2024-07-19 output/paper/manuscript.qmd
Compiled to PDF to check errors: missing acronyms and some references.
Working on comments Reviewer 1 points 9 to 11.

format:
  docx:
    page-width: 8.3
  pdf:
    documentclass: article
    geometry:
      - top=30mm
      - left=20mm
      - heightrounded
    pdf-engine: lualatex
    keep-tex: false
    include-in-header:
      text: |
        \usepackage{lineno}
        \usepackage{rotating}
        \usepackage{float}
        \usepackage{typearea}
        \usepackage{amsmath, amssymb, xcolor}
        \usepackage{pgf, tikz}
        \usetikzlibrary{external, trees, backgrounds, swigs, calc, positioning, shapes, arrows, arrows.meta, shapes.arrows, shapes.geometric, shapes.multipart, decorations.pathmorphing, decorations.markings}
    include-before-body:
      text: |
        \linenumbers

@@

filters:
  - acronyms
  - ../../../../all_phd/_extensions/wordcount/wordcount.lua

acronyms:
  fromfile: ../../../../all_phd/acronyms.yml
  insert_links: false
  insert_loa: false
  sorting: alphabetical

@@

For each research question, defined by a specific type of exposure and outcome, the minimal set of covariates for inclusion in the analyses was selected on the basis of a {{< acr dag >}} built with `DAGitty` [@TextorvanderZanderGilthorpe:2016] and `ggdag` [@Barrett:2023]. The sets of covariates were selected to estimate the total effect of the exposure on the outcome. For effect estimation of the {{< acr edcs >}} on glucocorticosteroids and of glucocorticosteroids on {{< acr hitrtse >}}, these sets were also sufficient to estimate direct effects. Sample-specific creatinine values were used to adjust for possible dilution effects. Further, each minimal adjustment set was *augmented* with precision covariates, defined as the set of parents variable of the outcome that are not parents of the exposure. Common confounders were cohort, ethnicity, sex, age, height, weight, and head circumference of the child, consumption of fish, fruit, vegetables, organic food, and fast food with a food frequency questionnaire, maternal tobacco consumption, family financial situation and affluence scale (FAS). Models for estimating the effects of {{< acr edcs >}} on {{< acr hitrtse >}} were further adjusted for child breastfeeding, prenatal maternal active and passive smoking, urine creatinine, child mood and rest before assessment, child neuropsychological diagnosis, marital status, season, and fasting time before assessment. Models for estimating the effects of {{< acr edcs >}} on glucocorticosteroids were further adjusted for urine creatinine, season, and fasting time before assessment. Models for estimating the effects of glucocorticosteroids on {{< acr hitrtse >}} were further adjusted for child breastfeeding, prenatal maternal active and passive smoking, marital status, {{< acr edcs >}}, urine creatinine, child mood and rest before assessment, and child neuropsychological diagnosis. The adjustment sets are provided in the Supplementary Material as text files compatible with DAGitty. Codebooks for the used covariates, by research question, are provided in Supplementary Tables [-@supptbl-codebook-1; -@supptbl-codebook-2; -@supptbl-codebook-3].

@@

@Tbl-hypothesis-1and2 presents the results of the difference between estimates of the {{< acr mcs >}} on the logarithmic scale for females and males, for the {{< acr edcs >}} on the glucocorticosteroids and {{< acr hitrtse >}}.
For {{< acr hitrtse >}}, significant differences were present for the phenol OXBE (`r format_res(heffects1, "oxbe", "HRT-SE")`) and the phthalate metabolite MbZP (`r format_res(heffects1, "mbzp", "HRT-SE")`).
For the glucocorticosteroids, significant differences were present across all three classes of {{< acr edcs >}} and for all outcomes. The largest differences were attributable to the phenol ETPA (corticosterone production, `r format_res(heffects2, "etpa", "corticosterone production")`) and the phthalate metabolite MEHP (cortisol production, (`r format_res(heffects2, "mehp", "cortisol production")`); cortisone production, (`r format_res(heffects2, "mehp", "cortisone production")`)). The forest plots of the individual {{< acr mcs >}} are presented in Supplementary Figures [-@suppfig-marginal-1sa] and [-@suppfig-marginal-2sa].

>>>>

format:
  docx:
    page-width: 8.3

@@

filters:
  - acronyms

acronyms:
  fromfile: acronyms.yml
  insert_links: false
  insert_loa: false
  sorting: alphabetical

@@

For each research question, defined by a specific type of exposure and outcome, the minimal set of covariates for inclusion in the analyses was selected on the basis of a {{< acr dag >}} built with `DAGitty` [@TextorvanderZanderGilthorpe:2016] and `ggdag` [@Barrett:2023]. The sets of covariates were selected to estimate the total effect of the exposure on the outcome. For effect estimation of the {{< acr edcs >}} on glucocorticosteroids and of glucocorticosteroids on {{< acr hitrtse >}}, these sets were also sufficient to estimate direct effects. Sample-specific creatinine values were used to adjust for possible dilution effects. Further, each minimal adjustment set was *augmented* with precision covariates, defined as the set of parents variable of the outcome that are not parents of the exposure. Common confounders were cohort, ethnicity, sex, age, height, weight, and head circumference of the child, consumption of fish, fruit, vegetables, organic food, and fast food with a food frequency questionnaire, maternal tobacco consumption, family financial situation and affluence scale (FAS). Models for estimating the effects of {{< acr edcs >}} on {{< acr hitrtse >}} were further adjusted for child breastfeeding, prenatal maternal active and passive smoking, urine creatinine, child mood and rest before assessment, child neuropsychological diagnosis, marital status, season, and fasting time before assessment. Models for estimating the effects of {{< acr edcs >}} on glucocorticosteroids were further adjusted for urine creatinine, season, and fasting time before assessment. Models for estimating the effects of glucocorticosteroids on {{< acr hitrtse >}} were further adjusted for child breastfeeding, prenatal maternal active and passive smoking, marital status, {{< acr edcs >}}, urine creatinine, child mood and rest before assessment, and child neuropsychological diagnosis. The adjustment sets are provided in the Supplementary Material as text files compatible with DAGitty. A simplified {{< acr swig >}} is provided in @fig-dag. Codebooks for the used covariates are provided in @supptbl-codebooks.

@@

@Tbl-hypothesis-1and2 presents the results of the difference between estimates of the {{< acr mcs >}} on the logarithmic scale for females and males, for the {{< acr edcs >}} on the glucocorticosteroids and {{< acr hitrtse >}}.
For {{< acr hitrtse >}}, significant differences were present for the phenol OXBE (`r format_res(heffects1, "oxbe", "HRT-SE")`) and the phthalate metabolite MbZP (`r format_res(heffects1, "mbzp", "HRT-SE")`).
For the glucocorticosteroids, significant differences were present across all three classes of {{< acr edcs >}} and for all outcomes. The largest differences were attributable to the phenol ETPA (corticosterone production, `r format_res(heffects2, "etpa", "corticosterone production")`) and the phthalate metabolite MEHP (cortisol production, (`r format_res(heffects2, "mehp", "cortisol production")`); cortisone production, (`r format_res(heffects2, "mehp", "cortisone production")`)). The individual {{< acr mcs >}} are presented in Supplementary Tables [-@supptbl-marginal-1sa], [-@supptbl-marginal-2sa] and [-@supptbl-marginal-3sa], and in Supplementary Figures [-@suppfig-marginal-1sa], [-@suppfig-marginal-2sa] and [-@suppfig-marginal-3sa].
